---
title: "A-Tailing"
output: html_notebook
params:
  num_samples: 4
  stopped: yes
---
```{r setup}
library(tidyverse)
```

If you are stopping here for a safe stopping point, resuspend the beads in 7 ul of 1x A Tailing buffer, otherwise, resuspend them in 15uL A-tailing master mix.  
```{r}
a_tailing <- tribble(~item, ~quantity,
                     "1x_a-tailing-buffer",ifelse(params$stopped == "yes", 7.5, 0),
                     "1x_a-tailing-master_mix",ifelse(params$stopped == "yes", 7.5, 15))
```

- Resuspend the beads in 15 ul 1X A-Tailing Buffer (see table above), cover the reaction and store at 4 oC for up to 24 hours. Do not freeze the samples as this will damage the AMPure® XP® beads. When ready, proceed to --A-Tailing after Safe Stopping Point--.

---

### A-Tailing
- A-Tailing is performed either directly after the 2nd Strand Synthesis and Marking Cleanup, or after the Safe Stopping Point, where beads were resuspended in 1X A-Tailing Buffer and stored at 4 °C for up to 24 hours. 

#### A-Tailing immediately
- Assemble the A-Tailing reaction as follows:

|Component|Volume|
|---------|------|
|Beads with dscDNA| --|
|A-Tailing Master Mix| 15 ul|
|--Total Volume--| --15 ul--|

- Mix thoroughly by pipetting up and down several times.
- Incubate the plate/tube using the following protocol:

|Step|Temp.|Duration|
|----|-----|--------|
|A-Tailing|30 °C|30 min|
|Enzyme inactivation|60 °C|30 min|
|HOLD|4 °C|∞|

- Proceed immediately to --Adapter Ligation--.

#### A-Tailing after safe stopping point
- To resume library preparation, combine the following reagents to perform A-Tailing:

|Component|Volume|
|---------|------|
|Beads with dscDNA (in 1X A-Tailing Buffer)| 7.5 ul |
|A-Tailing Master Mix after Safe Stopping Point| 7.5 ul|
|--Total Volume--| --15 ul--|

- Mix thoroughly by pipetting up and down several times.
- Incubate the plate/tube using the following protocol:

|Step|Temp.|Duration|
|----|-----|--------|
|A-Tailing|30 °C|30 min|
|Enzyme inactivation|60 °C|30 min|
|HOLD|4 °C|∞|

- Proceed immediately to --Adapter Ligation--.
