---
title: "Making Working Adapters/Oligos/Primers"
output: 
  html_notebook:
    toc: TRUE
params:
  initial_conc_uM: 200
  working_conc_uM: 10
  final_vol_ul: 100
  nmol_oligo: 9
  adpater_stock: "r_p2_primer_illumina_index_stock_ul"
---
```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
```

## Order oligos
- Log on to the marketplace at my.rutgers and click on IDT on the home screen.  ~~Go to [IDT website](www.idtdna.com), log in with the usualy pinsky credentials (MRS has in contacts).~~ View past orders. Click on the order and then hit the plus sign to see which oligos were purchased. Hit edit to update or change the oligo (like dropping the biotinylation from P2.2 or adding a phosphotioated bonds).  
- To order new oligos that have never been used before, sign in and click on custom DNA or RNA oligos, click on "Order Now" which should take you to an "Oligo Entry" page, and copy and paste your sequence. If you have more than one, click on "Bulk Input" to upload a file or copy and paste multiple sequences at once.  This works best if you choose comma delimited so you can see where new lines vs. spaces are.  Then click "add to order".  
**Double check the guaranteed nmole that you will be receiving.**

## Rehydrate oligos
IDT says to multiply # of nmoles on vial by 10 and add the answer uL of buffer to bring concentration to 100uM.  We want 200uM so we are multiplying by 5.

Add `r params$nmol_oligo / params$initial_conc_uM * 1000` ul buffer.

The forumua is:

$x L = \frac{mol\,dried\,oligo} {M\,rehydrated\, oligo}$


- Spin dry tubes 30 sec @ 20,000g
- Add TE buffer
- Vortex 15 sec
- Incubate at 4˚C overnight

This recipe is the same for ddRADSeq and EecSeq protocols
```{r working, include=FALSE}
working_guide <- tribble(~item, ~initial_conc, ~final_conc, 
                   params$adpater_stock, params$initial_conc_uM, params$working_conc_uM,
                   "ph2o_ul", 0, 0) %>% 
  mutate(final_vol_ul = params$final_vol_ul,
         initial_vol_ul = final_conc * final_vol_ul/initial_conc,
         initial_vol_ul = ifelse(item == "ph2o_ul",final_vol_ul - sum(initial_vol_ul, na.rm = T), initial_vol_ul))
 
working <- working_guide %>% 
  select(item, initial_vol_ul) %>% 
  rename(quantity = initial_vol_ul) %>% 
  mutate(item = ifelse(item == "ph2o_ul", "ph2o_ml", item), 
         quantity = ifelse(item == "ph2o_ml", quantity/1000, quantity)) %>% 
  bind_rows(tribble(~item, ~quantity,
                    "tips_10", 1,
                    "tips_100", 1,
                    "tubes_1.7ml", 1))
```

```{r echo=FALSE}
kable(working_guide) %>%
  kable_styling("bordered") %>%
  add_header_above(c("working stock solution" = 5))
```

```{r include=FALSE}
rm(working_guide)
```
