---
title: "Ampure SPRI cleanup"
output: html_notebook
params:
  sample_vol_ul: 25
  num_samples: 1
  ampure_vol_ul: 25
---
```{r setup, include=FALSE}
library(tidyverse)
```

# Perform a 1X SPRI® cleanup by combining the following:
```{r}
ampure <- tribble(~item, ~quantity,
                  # make room for small or large numbers of samples
                  ifelse(params$num_samples <20, "tubes_200ul", "plates"), ifelse(params$num_samples < 20, params$num_samples * 2, ceiling(params$num_samples/96) * 2),
                  "tips_100", params$num_samples * 5, 
                  "tips_1000", params$num_samples * 5, 
                  "ethanol_ul", params$num_samples * 400, 
                  "10_mM_Tris-HCl_pH8_ul", params$num_samples * 25)
```


* Thoroughly resuspend the beads by pipetting up and down multiple times.
* Incubate the plate/tube at room temperature for 5 – 15 min to allow the DNA to bind to the beads.
* Place the plate/tube on a magnet to capture the beads. Incubate until the liquid is clear.
* Carefully remove and discard 115 μl of supernatant.
* Keeping the plate/tube on the magnet, add 200 μl of 80% ethanol.
* Incubate the plate/tube at room temperature for ≥30 sec.
* Carefully remove and discard the ethanol.
* Keeping the plate/tube on the magnet, add 200 μl of 80% ethanol.
* Incubate the plate/tube at room temperature for ≥30 sec.
* Carefully remove and discard the ethanol. Try to remove all residual ethanol without disturbing the beads.
* Dry the beads at room temperature, until all of the ethanol has evaporated. **Caution: over-drying the beads may result in dramatic yield loss.**
* Remove the plate/tube from the magnet.
* Thoroughly resuspend the beads in 25 μl of 10 mM Tris-HCl (pH 8.0). 
* Incubate the plate/tube at room temperature for 2 min to allow the DNA to elute off the beads.
* Place the plate/tube on a magnet to capture the beads. Incubate until the liquid is clear.
* Transfer 24 μl of the clear supernatant to a new plate/tube and proceed to next step.
