---
title: "Anneal Adapters"
output: html_notebook
params:
  initial_conc_uM: 100
  p1_adapter_name: "sai_universal_stock_adapter_ul"
  p2_adapter_name: "sai_indexed_stock_adapter_ul"
---

```{r include=FALSE}
library(tidyverse)
```

```{r include=FALSE}
anneal <- tribble(~item, ~quantity,
                  params$p1_adapter_name, 40,
                  params$p2_adapter_name, 40, 
                  "10x_annealing_stock_ul", 10,
                  "ph2o_ml", 0.010, 
                  "tube_screw_0.5ml", 1,
                  "tips_100", 4)

```

Single-stranded oligos need to be annealed with their appropriate partner before ligation. 
 
- To create a duplex adapter, combine in a 1:1 ratio in working strength [annealing buffer](https://pinskylab.github.io/laboratory/protocols/annealing_stock.nb.html) (final buffer concentration 1x) for a total annealed adapter concentration of 40uM (for example, if purchased oligos are resuspended to an initial concentration of 100uM, use 40uL adapter part A, 40ul Adapter part B, 10ul 10x annealing buffer and 10ul nuclease-free water). 


**2.   In a thermocyler, incubate at 97.5°C for 2.5 minutes, and then cool at a rate of not greater than 3°C per minute until the solution reaches a temperature of 21°C. Hold at 4°C.**

3.   Prepare final working strength concentrations of annealed adapters from this annealed stock (the appropriate working stock dilution for your experiment can be determined from the chart below). For convenience, it is possible to store the adapters at 4°C while in active use.  
