---
title: "Bioanalyzer"
output: html_notebook
params:
  num_samples: 1
---
# Enter the number of libraries to be analyzed in the chip above
Contact Nicole Wagner (new39@rutgers.edu) to schedule a bioanalyzer run and get instructions from her regarding how to give her a sample of your library.

The High Sensitivity Chip can handle 12 libraries up to 7 kilobases.  If your library is larger, contact Nicole for pricing of a different chip.

```{r setup}
library(tidyverse)
```

```{r}
bioanalyzer <- tribble(~item, ~quantity,
                       "bioanalyzer_wells", params$num_samples)
```


# **Nicole does all of the steps below**
Note that a Tapestation could also be used at this step.

#### Bioanalyzer Instructions

#### Prepping the Gel (use filtered gel within 4 weeks)

- Pipette 550uL gel matrix (red) into a spin filter
- Centrifuge at 1500g for 10 mins at room temp
- Aliquot 65uL filtered gel into 0.5mL RNase-free tubes

#### Prepping the Gel-Dye Mix

- Allow RNA dye (blue) to equilibrate to room temp for 30 mins (in the dark)
- Vortex RNA dye (blue) for 10 secs and spin down
- Add 1 uL RNA dye into a 65uL aliquot of filtered gel
- Vortex solution
- Spin gel-dye mix at 13,000g for 10 mins at room temp

#### Turn on Laptop/Bioanalyzer (use gel-dye mix within 1 day)

- Select correct icon
- Select correct program (RNA, DNA, protein)

#### Cleaning the Electrodes

- Slowly fill one of the wells of the electrode cleaner with 350uL RNase-free water
- Place the electrode cleaner in the bioanalyzer
- Close the lid and let sit for 5 mins
- Upon removing the electrode cleaner, keep lid open for at least 30 secs (to allow for evaporation)

#### Loading the Gel-Dye Mix

- Put a new chip on the chip priming station
- Pipette 9uL of gel-dye mix in the well marked G (white G on black background)
- Position the plunger at 1mL
- Close the chip priming station (MAKE SURE IT CLICKS)
- Press plunger down until it’s held by the clip
- Wait exactly 30 secs and release clip
- Wait exactly 5 secs and slowly pull plunger back to 1mL
- Open chip priming station and pipette 9uL gel-dye mix into wells marked G (black G on grey background)
- Discard remaining gel-dye mix

#### Loading the Conditioning Solution and Marker 

- Pipette 9uL of RNA conditioning solution (white) into well marked CS
- Pipette 5uL of RNA marker (green) in all 11 sample wells and the ladder well

#### Loading the Diluted Ladder and Samples

- Pipette 1uL ladder into the well marked ladder
- Pipette 1uL of sample into each of the 11 sample wells (pipette 1uL RNA marker (green) into each unused sample well)
- Vortex chip for 1 min at 2400 rpm
- Run the chip in the bioanalyzer within 5 mins (RNA assay)

--CLEAN THE ELECTRODES AGAIN AFTER EVERY RUN--
