---
title: "Biotin Labeling"
output: html_notebook
params:
  num_samples: 1
---
```{r setup}
library(tidyverse)
```

```{r}
biotin <- tribble(~item, ~quantity,
                  "tips_100", 4,
                  "tips_10", 6, 
                  "ph2o_ml", 0.023,
                  "tubes_1.7ml", 2,
                  "decalabel_biotin_klenow_ul", 2) %>% 
  mutate(quantity = quantity * params$num_samples)
```

### Materials needed

| Reagent                                | Producer         | Catalog #      |            
|----------------------------------------|------------------|-----------|
|DecaLabel™ Biotin DNA Labeling Kit |Thermo Scientific | FERK0651   |                     

### Procedure

- Add the following components into 1.5 ml microcentrifuge tube:

|Component|Volume|
|---------|------|
|RE and MBN treated DSN Library| 20 ul|
|Decanucleotide in 5X Reaction Buffer| 10 ul|
| Water, nuclease-free| 14 ul|
|--Total Volume--| --44 ul--|

- Vortex the tube and spin down in a microcentrifuge for 3-5 s
- Incubate the tube in a boiling water bath for 5-10 min and cool it on ice. Spin down quickly.
- Add the following components in the same tube:

|Component|Volume|
|---------|------|
|Biotin Labeling Mix| 5 ul|
|Klenow fragment, exo– (5 u)| 1 ul|
|--Total Volume--| --50 ul--|

- Shake the tube and spin down in a microcentrifuge for 3-5 s. 
- Incubate for 12-20 hours at 37°C. 

---
#### Optional:Control reaction
- Add the following components into 1.5 ml microcentrifuge tube:

|Component|Volume|
|---------|------|
|Control Template, 10 ng/ul | 25 ul|
|Decanucleotide in 5X Reaction Buffer| 10 ul|
| Water, nuclease-free| 9 ul|
|--Total Volume--| --44 ul--|

- Vortex the tube and spin down in a microcentrifuge for 3-5 s
- Incubate the tube in a boiling water bath for 5-10 min and cool it on ice. Spin down quickly.
- Add the following components in the same tube:

|Component|Volume|
|---------|------|
|Biotin Labeling Mix| 5 ul|
|Klenow fragment, exo– (5 u)| 1 ul|
|--Total Volume--| --50 ul--|
