---
title: "DSN Preparation"
subtitle: "updated `r Sys.Date()`"
output: html_notebook
params:
  num_rna: 22
  num_lib: 1
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
```


```{r}
dsn_norm <- tribble(~item, ~quantity,
               "tips_100", params$num_rna,
               "tips_100", 5,
               "tips_10", 3,
               "tubes_200ul", 3,
               "tubes_1.7ml", 1,
               "tips_1000", 3,
               "1_M_HEPES_ul", 200,
               "5_M_nacl_ul", 400,
               "ph2o_ml", 0.4, 
               "dsn_kit", 1,
               "tube_screw_1.5ml", 1) %>% 
  mutate(quantity = quantity * params$num_lib)

kable(dsn_norm) %>% 
  kable_styling("bordered") %>% 
  add_header_above(c("Supplies" = 2))
```

DSN normalization is critical ensuring an even distribution of coverage across probes.  There are a genes that are highly expressed in all cells and DSN normalization helps to remove these high abundance probes and transcripts.  

#### DSN needs to be properly diluted and should be tested for activity levels before proceeding

#### The protocol below was taken from [Illumina's recommendations](https://github.com/pinskylab/laboratory/blob/master/documentation/dsn_normalization_sampleprep_application_note_15014673_c.pdf)
#### Reagents

| Reagent| Supplier|
|----------|--------------|
|1 M HEPES buffer solution|Invitrogen, part # 15630‐080 |
|5 M NaCl solution|Ambion, part # AM9760G|
|KAPA HiFi HotStart PCR kit with dNTPs|Kapa, part #KK2502|
|Strip tubes|General lab supplier|
|DSN Kit|Evrogen, part # EA001 Sigma Aldrich, part # E7023|
|Ethanol 200 proof (absolute) for molecular biology (500 ml)|AB, part # 4333764F|
|PCR Primer PE 1.0|Included in Kapa stranded mRNA kit|
|PCR Primer PE 2.0|Included in Kapa stranded mRNA kit|
|SPRI beads|Agencourt AMPure, part # 29152; KAPA Pure Beads, part #KK8000|
|Nuclease-free water|General lab supplier|

#### Equipment
- Thermocycler
- Magentic stand compatible with strip tubes

#### Things that can be done ahead of time
1. Pool individual RNA libraries in equal quantities to create a single pool of 500 ng.  This pool will now be called the sample. *For example pool 125 ng each of four individual libraries.*  

2. Create a 4X hybridization solution and store in a screw cap tube for future use at -15˚C to -25˚C:
*This recipe has been cut in half to fit the 0.5mL screw cap tubes we currently have in stock, it normally is for 1000 uL, this recipe is prepared for `r params$num_lib` library(ies) pooled from `r params$num_rna` sample(s).*
```{r}
kable(tribble(~Reagent, ~Volume_ul,
              "1 M HEPES buffer solution",100,
              "5 M NaCl solution",200,
              "Nuclease free water", 200,
              "Total Volume Per sample", 500) %>% 
        mutate(Volume_ul = Volume_ul*params$num_lib)) %>% 
  kable_styling("bordered")
```

3. Set one side thermocycler or heat block to hold at 68˚C and pre-heat it.

#### Procedure

1. Prepare the following reaction mix in a separate, sterile, nuclease‐free 200 ul PCR tube on ice for each sample to be normalized.

|Component|Volume|
|---------|------|
|Sample library (500 ng)| 13.5 ul|
|4X Hybridization buffer| 4.5 ull|
|Total Volume Per Sample|18 ul|

* NOTE: The sample library volume is optimized for the removal of rRNA from a non‐poly‐A selected library. For a reduction of highly expressed transcripts from a poly‐A selected library, more sample library may be needed for optimal results.*

2. Gently pipette the entire volume up and down 10 times, then centrifuge briefly to mix.  
3. Transfer the entire volume of reaction mix directly to the bottom of a new, sterile, nuclease‐free 200 ul PCR tube, using a pipette. Do not let the sample contact the side of the tube during the process.  

4. Incubate the reaction mix tube on the thermal cycler using the following PCR cycling conditions:

|Step|Temp|Duration|
|----|----|--------|
|Initial denaturation|98 °C|2 min|
|Treatment|68 °C|5 hours|

**Caution: Following incubation, keep the thermal cycler lid closed and the temperature held at 68°C. Do not remove the reaction mix tube from thermal cycler prior to and during DSN treatment.**
