---
title: "DSN PCR"
subtitle: "updated `r Sys.Date()`"
output: html_notebook
params:
  num_samples: 1
---
```{r setup}
library(tidyverse)
```

```{r}
dsn_pcr <- tribble(~item, ~quantity,
                   "kapa_hifi_hotstart_readymix_ul", 25,
                   "kapa_10x_library_amp_master_mix_ul", 5,
                   "tubes_200ul", 3,
                   "tips_10", 1,
                   "tips_100", 3
                   ) %>% 
  mutate(quantity = quantity * params$num_samples)
```

#### The protocol below was taken from [Illumina's recommendations](https://github.com/pinskylab/laboratory/blob/master/documentation/dsn_normalization_sampleprep_application_note_15014673_c.pdf)

### PCR Enrichment of DSN Normalized Library

1. Prepare the following PCR reaction mix in a separate, sterile, nuclease‐free tube for each of the samples being prepared.  Mix well by pipetting up and down several times.  

|Component|Volume|
|---------|------|
|DSN Treated Library | 30 ul|
|2X KAPA HiFi HotStart ReadyMix| 25 ul|
|10X KAPA Library Amplification Primer Mix| 5 ul|
|**Total Volume per sample**|**50 ul**|


2. Amplify the library using the following thermal cycling protocol:  


|Step|Temp|Duration|Cycles|
|----|----|--------|------|
|Initial denaturation|98 °C|45 sec|1|
|Denaturation|98 °C|15 sec|12|
|Annealing|60 °C|30 sec|12|
|Extension|72 °C|30 sec|12|
|Final Extension|72 °C|5 min|1|
|Hold|10 °C | ∞|1|

