---
title: "End Repair"
subtitle: "updated `r Sys.Date()`"
output: html_notebook
params:
  num_samples: 160
---

[Manufacturer's instructions](https://github.com/pinskylab/laboratory/blob/master/documentation/kapa-hyper-plus-dna.pdf)

```{r setup, include=FALSE}
library(tidyverse)
```

```{r}
repair <- tribble(~item, ~quantity,
                  "tips_10", 2,
                  )
```

2.1 In the same plate/tube(s) in which enzymatic fragmentation was performed, assemble each end repair and A-tailing reaction as follows:

|Component|Volume|1/2 rxn|
|----|----|----|
|Fragmented, double-stranded DNA|50 μL| 25 μL|
End Repair & A-Tailing Buffer*|7 μL| 3.5 μL|
|End Repair & A-Tailing Enzyme Mix*|3 μL| 1.5 μL|
|**Total volume:**|**60 μL**|**30 μL**|

*The buffer and enzyme mix should preferably be pre-mixed and added in a single pipetting step. Premixes are stable for ≤24 hrs at room temperature, for ≤3 days at 4°C, and for ≤4 weeks at -20°C.*

2.2 Vortex gently and spin down briefly. Return the reaction plate/tube(s) to ice. Proceed immediately to the next step.


2.3 Incubate in a thermocycler programmed as outlined below. A heated lid is required for this step. If possible, set the temperature of the heated lid to ~85°C (instead of the usual 105°C).

|Step|Temp|Time|
|----|----|--------|
|End repair and A-tailing 2|65 °C|30 min|
|Hold|4°C | ∞|

*- Both the fragmentation and end repair enzymes are inactivated at 65°C. When reactions are set up according to recommendations, additional fragmentation should be negligible. The brief period of end repair is sufficient for enzymatically fragmented DNA.*

*- If proceeding to the adapter ligation reaction setup without any delay, the reaction may be cooled to 20°C instead of 4°C.*

