---
title: "KAPA SPRI cleanup"
params:
  sample_vol_ul: 55
  num_samples: 160
  bead_vol_ul: 44
---
```{r setup, include=FALSE}
library(tidyverse)
```

# Perform a 1X SPRI® cleanup by combining the following:
```{r}
kapa_clean <- tribble(~item, ~quantity,
                  # make room for small or large numbers of samples
                  ifelse(params$num_samples <20, "tubes_200ul", "plates"), ifelse(params$num_samples < 20, params$num_samples, ceiling(params$num_samples/96)),
                  "tips_10", params$num_samples * 1,
                  "tips_100", params$num_samples * 2, 
                  "tips_1000", params$num_samples * 5, 
                  "ethanol_ul", params$num_samples * 400, 
                  "10_mM_Tris-HCl_pH8_ul", params$num_samples * 25, 
                  "kapa_beads_ul",params$num_samples * params$bead_vol_ul)
```

- In the same plate/tube(s), perform a 0.8X bead- based cleanup by combining the following:

|Component|Volume|
|---------|------|
|Adapter ligation reaction product| 55 μl|
|KAPA Pure Beads | 44 μl|
|--Total Volume--| --99 μl--| 

- Mix thoroughly by vortexing and/or pipetting up and down multiple times.
- Incubate the plate/tube(s) at room temperature for 5 – 15 min to bind DNA to the beads.
- Place the plate/tube(s) on a magnet to capture the beads. Incubate until the liquid is clear.
- Carefully remove and discard the supernatant.
- Keeping the plate/tube(s) on the magnet, add 200 μL of 80% ethanol.
- Incubate the plate/tube(s) on the magnet at room temperature for ≥30 sec.
- Carefully remove and discard the ethanol.
- Keeping the plate/tube(s) on the magnet, add 200 μL of 80% ethanol.
- Incubate the plate/tube(s) on the magnet at room temperature for ≥30 sec.
- Carefully remove and discard the ethanol. Try to remove all residual ethanol without disturbing the beads.
- Dry the beads at room temperature for 3 – 5 min, or until all of the ethanol has evaporated. -Caution: over-drying the beads may result in reduced yield.-
- Remove the plate/tube(s) from the magnet.
- Thoroughly resuspend the beads in in 12.5 μL of elution buffer (10 mM Tris-HCl, pH 8.0 – 8.5)
- Incubate the plate/tube(s) at room temperature for 2 min to elute DNA off the beads.
- Place the plate/tube(s) on a magnet to capture the beads. Incubate until the liquid is clear.
- Transfer 11 μL of supernatant to a new plate/tube(s):
