---
title: "Library Amplification"
output: html_notebook
params:
  num_seq_lib: 1
---
```{r setup, include = FALSE}
library(tidyverse)
library(knitr)
```

### Library Amplification

- Assemble each library amplification reaction as follows per pool:
```{r}
amp <- tribble(~item, ~quantity,
                 "kapa_hifi_hotstart_readymix_ul", 12.5 * params$num_seq_lib,
                 "f_p1_primer_universal_working_ul", 1.25 * params$num_seq_lib,
                 "r_p2_primer_illumina_index_working_ul", 1.25 * params$num_seq_lib, 
                 "tips_100", 1,
                 "tips_10", 3 * params$num_seq_lib,
                 # make room for small or large numbers of samples
                  ifelse(params$num_seq_lib <20, "tubes_200ul", "plates"), ifelse(params$num_seq_lib < 20, params$num_seq_lib * 2, ceiling(params$num_seq_lib/96) * 2))


```


|Component|Volume|
|---------|------|
|KAPA HiFi HotStart ReadyMix (2X) | 12.5 μl|
|PCR1_P5 Primer (Universal)  | 1.25 μl|
|PCR2 Primer (Indexed)  | 1.25 μl|
|Adapter-ligated library pool| 10.0 μl|
|--Total Volume--| --25 μl--| 

- Calculate number of cycles needed based on previous quants

|Amount of adapter-ligated DNA in amplification reaction| Number of cycles required to generate 1 μg of library DNA| 
|--------------|-------------------------------------------|
|500 ng|1-2|
|100 ng|3-4|
|50 ng|5-6|
|25 ng|7-8|
|10 ng|8-9|
|5 ng|11-12|
|1 ng|12-13|

### For Puritz and Lotterhos (2017), 6 PCR cycles were used.

- Mix thoroughly and centrifuge briefly.
- Amplify using the following cycling protocol:

|Step|Temp|Duration|Cycles|
|----|----|--------|------|
|Initial denaturation|98 °C|45 sec|1|
|Denaturation|98 °C|15 sec|X|
|Annealing-|60 °C|30 sec|X|
|Extension|72 °C|30 sec|X|
|Final Extension|72 °C|1 min|1|
|Hold|4 °C | ∞|1|

- Proceed immediately to the next step
