---
title: "Ligation - EecSeq"
params:
  num_samples: 160
---
```{r setup, include = FALSE}
library(tidyverse)
```


### Adapter ligation

```{r}
ligation <- tribble(~item, ~quantity, 
                    "p1_adapter_eecseq", 1.25,
                    "p2_adapter_eecseq", 1.25,
                    "ph2o_ml", 0.0025,
                    "dna_ligase_ul", 5,
                    "tips_10",4, 
                    "tips_100", 1) %>% 
  mutate(quantity = quantity * params$num_samples) %>% 
  bind_rows(tribble(~item, ~quantity,
                    ifelse(params$num_samples <20, "tubes_200ul", "plates"), ifelse(params$num_samples < 20, params$num_samples, ceiling(params$num_samples/96))))
```

- Dilute adapter stocks to the appropriate concentration, as outlined below:

|Fragmented DNA| Adapter stock concentration | Adapter:insert molar ratio|
|--------------|-----------------------------|---------------------------|
| 1 μg|15 μM|10:1|
|500 ng|15 μM|20:1|
|250 ng|15 μM|40:1|
|100 ng|15 μM|100:1|
|50 ng|15 μM|200:1|
|25 ng|15 μM|200:1|
|10 ng|15 μM|200:1|
|5 ng|15 μM|200:1|
|2.5 ng|15 μM|200:1|
|1 ng|15 μM|200:1|

--For Puritz and Lotterhos (2017), a working stock of 40 μM was used, leading to a final adapter:insert molar ratio of ~ 50:1.--

- In the same plate/tube(s) in which end repair and A-tailing was performed, assemble each adapter ligation reaction as follows:

|Component|Volume|
|---------|------|
|End repair and A-tailing reaction product| 30 μl|
|P1 Adapter stock (concentration as required) --Barcode containing oligo-- | 1.25 μl|
|P2 Adapter stock (concentration as required) | 1.25 μl|
|PCR-grade water-| 2.5 μl |
|Ligation Buffer-| 15 μl |
|DNA ligase-| 5 μl |
|--Total Volume--| --55 μl--|

- Notes
  - --Each P1 adapter has a unique barcode-- these barcodes can be combined with Illumina indices (added later via PCR) for high levels of multiplexing. 
  - The water, buffer and ligase enzyme should preferably be premixed and added in a single pipetting step. Premixes are stable for ≤24 hrs at room temperature, for ≤3 days at 4°C, and for ≤4 weeks at -20°C.
  
- Mix thoroughly and centrifuge briefly.
- Incubate at 20°C for 60 min.
  - Note: to achieve higher conversion rates and library yields, particularly for low-input samples, consider increasing the ligation time—to a maximum of 4 hrs at 20°C, or overnight at 4°C. Please note that longer ligation times may lead to increased levels of adapter-dimer. Adapter concentrations may have to be optimized if ligation times are extended significantly.
- Proceed immediately to the next step.
