---
title: "mRNA Capture"
output: html_notebook
params:
  num_samples: 4
---
```{r setup, include=FALSE}
library(tidyverse)
```

```{r}
mrna_capture <- tribble(~item, ~quantity,
                     "kapa_mrna_capture_beads_ul", 26.25* params$num_samples,
                     "tubes_200ul", 1, 
                     "tips_100", 2,
                     "kapa_mrna_bead_binding_buffer_ul", 26.25* params$num_samples*2,
                   "tubes_200ul", params$num_samples,
                   "tips_100", 8 * params$num_samples,
                   "kapa_fragment_elute_buffer_ul", 5.5 * params$num_samples, 
                   "kapa_mrna_bead_wash_ul", 200*params$num_samples, 
                   "ph2o_ml", 0.025 * params$num_samples,
                   "kapa_mrna_bead_binding_buffer_ul", 25 * params$num_samples)
```

#### Adapter concentration will vary depending on overall RNA yield, see table below:
|Quantity of starting material|Adapter stock concentration|Adapter concentration in ligation reaction|
|----|----|-----|
|100 – 250 ng|140 nM| 10 nM|
|251 – 500 ng|350 nM|25 nM|
|501 – 2000 ng|700 nM|50 nM|
|2001 – 4000 ng|1400 nM|100 nM|

For Puritz and Lotterhos 2017, we used 4000 ng starting RNA, but because of difficulties assessing and quantifying molluscan RNA, we chose to use a 700 nM working stock with a final reaction concentration of 50 nM.

--Procedure:--
- Prepare mastermixes for number of libraries (individual RNA extractions)
  - See tables in manual and tables below for guidelines (We are using 1/2 reactions):

|Component| Total Volume Needed for 4 RXNs (Includes 20% excess)|
|---------|--------------------|
|--1st Strand Synthesis Master Mix:--|--|
|1st Strand Synthesis Buffer|22 μl|
|KAPA Script|2 μl|
|--Total Master Mix Volume--| --24 μl--|
|--Final reaction composition:--|--|
|1st Strand Synthesis Master Mix| 5 μl|
|Fragmented, primed RNA|10 μl| 
|--Total Reaction Volume--| --15 μl--|

|Component| Total Volume Needed for 4 RXNs (Includes 10% excess)|
|---------|--------------------|
|--2nd Strand Synthesis and Marking Master Mix:--|--|
|2nd Strand Marking Buffer|62 μl|
|2nd Strand Synthesis Enzyme Mix|4 μl|
|--Total Master Mix Volume--| --66 μl--|
|--Final reaction composition:--|--|
|2nd Strand Synthesis and Marking Master Mix| 15 μl|
|Fragmented, primed RNA|15 μl| 
|--Total Reaction Volume--| --30 μl--|

|Component| Total Volume Needed for 4 RXNs (Includes 10% excess)|
|---------|--------------------|
|--A-Tailing Master Mix:--|--|
|Water|52.8 μl|
|10X KAPA A-Tailing Buffer|6.6 μl|
|KAPA A-Tailing Enzyme|6.6 μl|
|--Total Master Mix Volume--| --66 μl--|
|--Resuspend beads in a volume of:--|-- 15 μl-- |

|Component| Total Volume Needed for 4 RXNs (Includes 10% excess)|
|---------|--------------------|
|--Adapter Ligation Master Mix:--|--|
|Water|35.2 μl|
|5X KAPA Ligation Buffer| 30.8μl|
|KAPA T4 DNA Ligase|11 μl|
|--Total Master Mix Volume--| --77 μl--|
|--Final reaction composition:--|--|
|Beads with A-tailed DNA|15 μl|
|Adapter Ligation Master Mix|17.5 μl|
|Adapter (350 nM – 1400 nM, as appropriate)|2.5 μl|
|--Total Reaction Volume--| --35 μl--|

|Component| Total Volume Needed for 4 RXNs (Includes 10% excess)|
|---------|--------------------|
|--Library Amplification Master Mix:--|--|
|2X KAPA HiFi HotStart ReadyMix|55 μl|
|10X KAPA Library Amplication Primer Mix|11 μl|
|--Total Master Mix Volume--| --66 μl--|
|--Final reaction composition:--|--|
|Adapter-ligated library DNA|10 μl|
|Library Amplification Master Mix|15 μl|
|Balance of water (if required)|5 μl|
|--Total Reaction Volume--| --30 μl--|

- Before mRNA capture beads can be used they must be washed with mRNA Bead Binding Buffer
	- Resuspend beads thoroughly by gentle pipetting or vortexing
	- For each library to be prepared, transfer 26.25 uL of the resuspended mRNA Capture beads into an appropriate tube
		- Up to 48 libraries (1,260 uL) can be washed in a single tube
	- Place the tube on a magnet holder and incubate at room temperature until solution is clear.
	- Discard supernatant and replace with an equal volume of mRNA Bead Binding Buffer.
	- Remove tube from magent and again resuspend the beads.
	- Place the tube on a magnet holder and incubate at room temperature until solution is clear.
	- Discard supernatant and replace with an equal volume of mRNA Bead Binding Buffer.
	- Remove tube from magent and again resuspend the beads.

- Combine the following for each RNA sample to be captured:

|Component|Volume|
|---------|------|
|RNA sample (in RNase-free water)| 25 μl|
|KAPA mRNA Capture Beads| 25 μl|
|--Total Volume--| --50 μl--|

- Mix thoroughly by gently pipetting up and down several times.
- Place the plate/tube in a thermal cycler and carry out the 1st mRNA capture program as follows:

|Step|Temp.|Duration|
|----|-----|--------|
|1st mRNA capture|65 °C|2 min|
|Cool|20 °C|5 min|


- Place the plate/tube containing the mixture of KAPA mRNA Capture Beads and RNA on a magnet and incubate at room temperature until the solution is clear. Remove and discard the supernatant.
- Remove the plate/tube from the magnet and resuspend thoroughly in 100 μl of KAPA mRNA Bead Wash Buffer by pipetting up and down several times.
- Place the plate/tube on the magnet and incubate at room temperature until the solution is clear. Remove and discard the supernatant.
- Resuspend the beads in 25 μl of RNase-free water.
- Place the plate/tube in a thermal cycler and carry out the 2nd mRNA capture program as follows:

|Step|Temp.|Duration|
|----|-----|--------|
|2nd mRNA capture|70 °C|2 min|
|Cool|20 °C|5 min|

- Add 25 μl of KAPA Bead Binding Buffer to the mixture of KAPA mRNA Capture Beads and RNA and mix thoroughly by gently pipetting up and down several times.
- Incubate the plate/tube at 20 °C for 5 min.
- Place the plate/tube on the magnet and incubate at room temperature until the solution is clear. Remove and discard the supernatant.
- Remove the beads from the magnet and resuspend in 100 μl of KAPA mRNA Bead Wash Buffer by pipetting up and down several times.
- Place the plate/tube on the magnet and incubate at room temperature until the solution is clear. Remove and discard the entire volume of supernatant.

### mRNA Elution, Fragmentation, and Priming
- Prepare the required volume of 1X Fragment, Prime and Elute Buffer as follows:

|Component|Volume per sample|
|---------|------|
|Water| 5.5 μl|
|Fragment, Prime and Elute Buffer (2X)| 5.5 μl|
|--Total Volume--| --11 μl--|

- Thoroughly resuspend the KAPA mRNA Capture Beads with captured mRNA in 11 μl of 1X Fragment, Prime and Elute Buffer.

---

### Safe Stopping Point
Resuspended beads with captured mRNA may be stored at 4 oC for up to 24 hours. Do not freeze the samples as this will damage the beads. When ready, proceed to step below.
