---
title: "Qubit - High Sensitivity"
output: html_notebook
params:
  num_samples: 1
---

# Enter the number of samples to be quantified above
```{r setup}
library(tidyverse)
```

## Supplies Needed
```{r}
qubit <- tribble(~item, ~quantity, 
                 "qubit_tubes", params$num_samples + 2,
                 "qubit_dye_ul", params$num_samples + 2,
                 "qubit_buffer_ul", 199 * (params$num_samples + 2),
                 "qubit_std1_ul", 10,
                 "qubit_std2_ul", 10, 
                 "tips_10", params$num_samples + 2,
                 "tips_100", 1,
                 "tips_1000", 2,
                 "tubes_1.7ml", 1)

```

## Procedure (Standard HS DNA protocol)
- Set up the required number of 0.5-mL tubes for standards and samples. The Qubit® RNA HS Assay requires 2 standards.
- Label the tube lids.
- Prepare the Qubit® working solution by diluting the Qubit® DNA HS Reagent 1:200 in Qubit® DNA HS Buffer. Use a clean plastic tube each time you prepare Qubit® working solution. **Do not mix the working solution in a glass container.**
- Add 190 μL of Qubit® working solution to each of the tubes used for standards.
- Add 10 μL of each Qubit® standard to the appropriate tube, then mix by vortexing 2–3 seconds. Be careful not to create bubbles.
- Add Qubit® working solution to individual assay tubes so that the final volume in each tube after adding sample is 200 μL.
- Add each sample to the assay tubes containing the correct volume of Qubit® working solution, then mix by vortexing 2–3 seconds. The final volume in each tube should be 200 μL.
- Allow all tubes to incubate at room temperature for 2 minutes.
- On the Home screen of the Qubit® 3.0 Fluorometer, press DNA, then select DNA: High Sensitivity as the assay type. The “Read standards” screen is displayed. Press Read Standards to proceed.
- Insert the tube containing Standard #1 into the sample chamber, close the lid, then press Read standard. When the reading is complete (~3 seconds), remove Standard #1.
- Insert the tube containing Standard #2 into the sample chamber, close the lid, then press Read standard. When the reading is complete, remove Standard #2.
- Press Run samples.
- On the assay screen, select the sample volume and units
- Insert a sample tube into the sample chamber, close the lid, then press Read tube. When the reading is complete (~3 seconds), remove the sample tube.
- Repeat step last step until all samples have been read
