---
title: "Remove Adapters from cDNA"
output: html_notebook
params:
  num_samples: 1
---
```{r}
library(tidyverse)
```

```{r}
remove <- tribble(~item, ~quantity,
                  "mung_bean_nuclease_ul", 0.5, 
                  "tips_10", 2,
                  "sali_hf_ul", 1, 
                  "tips_100", 2,
                  "ph2o_ml", 0.02275, 
                  "tubes_200ul", 1) %>% 
  mutate(quantity = quantity * params$num_samples)
```

#### Materials needed
| Reagent                                | Producer         | Catalog #      |            
|----------------------------------------|------------------|-----------|
|Mung Bean Nuclease| NEB| M0250S|
|SalI-HF| NEB| R3138T|
|Agencourt AMPure XP  |Beckman Coulter   | A63881|

### Procedure
1. Setup a restriction digest using 1 ug of DSN library

|Component|Volume|
|---------|------|
| 10X Cutsmart Buffer| 4 ul|
|SalI-HF Enzyme (100 units)| 1 ul |
| Molecular Grade H20| 22.75 ul|
| DSN Enriched Library| 12.25 ul|
|**Total Volume**| **40 ul**|

2. Incubate reactions in thermocycler at 37°C for at least 8 hours, prefereably 12-16 hours.

3. **Using the same tubes from the previous step** combine the following:

|Component|Volume|
|---------|------|
|Digested Library| 40 ul|
|10X Mung Bean Nuclease buffer| 4.5 ul|
|Mung Bean Nuclease (10 units per ul)| 0.5 ul|
|**Total Volume**| **45 ul**|

4. Incubate at 30°C for 30 minutes
