---
title: "RNA adapter ligation"
output: html_notebook
params:
  num_samples: 4
---
```{r setup}
library(tidyverse)
```

```{r}
rna_ligation <- tribble(~item, ~quantity,
                        "kapa_rna_adapter_ligation_master_mix_ul", 17.5, 
                        "tips_100", 2 * params$num_samples,
                        "tubes_200ul", 2 * params$num_samples, 
                        "sai_working_adapter_ul", 2.5 * params$num_samples)
```

#### Adapter concentration will vary depending on overall RNA yield, see table below:

|Quantity of starting material|Adapter stock concentration|Adapter concentration in ligation reaction|
|----|----|-----|
|100 – 250 ng|140 nM| 10 nM|
|251 – 500 ng|350 nM|25 nM|
|501 – 2000 ng|700 nM|50 nM|
|2001 – 4000 ng|1400 nM|100 nM|

For Puritz and Lotterhos 2017, we used 4000 ng starting RNA, but because of difficulties assessing and quantifying molluscan RNA, we chose to use a 700 nM working stock with a final reaction concentration of 50 nM.

#### This will be where we insert the custom adapters that are barcoded with RE sites

- Set up the adapter ligation reactions as follows:

|Component|Volume|
|---------|------|
|Beads with A-tailed DNA| 15 ul |
|Adapter Ligation Master Mix| 17.5 ul |
|--Adapters---| 2.5 ul|
|--Total Volume--| --35 ul--|


- Mix thoroughly by pipetting up and down several times to resuspend the beads.
- Incubate the plate/tube at 20 °C for 30 min.
- Proceed immediately to --1st Post-Ligation Cleanup--.
