---
title: "RNA extraction for EecSeq"
output: html_notebook
params:
  num_samples: 4
---

```{r setup, include = FALSE}
library(tidyverse)
```

## RNA Prep
```{r}
rna_extr <- tribble(~item, ~quantity,
                    "tri_reagent_ul", 1000,
                    "tubes_1.7ml", 3,
                    "bcp_ul", 100,
                    "tips_1000", 7,
                    "tips_100", 2,
                    "isopropanol_ul", 500, 
                    "ethanol_ul", 1000,
                    "1x_te_buffer_ul", 300
                    ) %>% 
  mutate(quantity = quantity* params$num_samples)
```


### Extract RNA from individuals to be used for probes
-Refer to manual during procedure (steps below are for notes and comments)-
### Using unmodified TRI Reagent Protocol [LINK](https://assets.thermofisher.com/TFS-Assets/LSG/manuals/9738M_D.pdf) (Below are summary steps)

#### Reagents and supplies
- Nuclease-free Water
- 1-bromo-3-chloropropane (BCP; recommended, e.g., MRC, Cat #BP 151), or chloroform without added isoamyl alcohol 100% 
- isopropanol, ACS grade or better
- 100% ethanol, ACS grade or better

#### Equipment
- Tissue homogenizer with replacable probes (such as Qiagen TissueRuptor®)
- Appropriately sized RNase-free centrifuge tubes with secure closures, compatible with phenol/chloroform (polypropyl- ene, or polyallomer), and capable of withstanding centrifugal forces of 12,000G
- Centrifuge capable of 12,000 x g

**Notes before starting**
- Place all pipettes, tips, and supplies inside the hood and expose to UV for 30 minutes. (Do not put your samples in the UV hood!)
- Prepare 75% ethanol by mixing 250 μL nuclease-free water with 750 μL 100% ethanol per mL of TRI Reagent solution to be used. Include 10% overage to ensure a sufficient volume.

#### Procedure
- Homogenize tissue samples in 10–20 volumes TRI Reagent solution. Homogenize cultured cells in 1 mL TRI Reagent solution per 5–10 x 10^6 cells, or per 10 cm^2 culture dish area.
- Incubate the homogenate for 5 min at room temp.
- (Optional) Centrifuge at 12,000 xg for 10 min at 4°C and transfer the supernatant to a fresh tube.
- Add 100 μL BCP per 1 mL of TRI Reagent solution used for homogenization. Alternatively, 200 μL of chloroform (without isoamyl alcohol) can be used in place of BCP.
	- Cap the tubes tightly and shake vigorously for 15 sec.
	- Incubate the mixture at room temperature for 5–15 min.
- Centrifuge at 12,000 x g for 10–15 min at 4°C, then transfer the aqueous phase to a fresh tube.
	- Centrifugation at temperatures > 8°C may cause some DNA to partition in the aqueous phase, which is undesireable.
	- Transfer the aqueous phase (colorless top layer) to a fresh tube.
- Add 500 μL of isopropanol per 1 mL of TRI Reagent solution, vortex for 5–10 sec, and incubate at room temp for 5–10 min.
- Centrifuge at 12,000 x g for 8 min at 4–25°C, and discard the supernatant.
	- Carefully remove the supernatant without disturbing the pellet.
		- Precipitated RNA forms a gel-like or white pellet on the side and bottom of the tube.
- Add 1 mL of 75% ethanol per 1 mL of TRI Reagent solution
- Centrifuge at 7,500 x g for 5 min at 4–25°C.  
- If the precipitated RNA floats or does not form a compact pellet, repeat the centrifugation at 12,000 x g for 5 min to consolidate the pellet at the bottom of the tube.
- Remove the ethanol wash without disturbing the pellet.
- Remove all residual ethanol by centrifuging again briefly and removing the ethanol that collects with a fine tip pipette. **Complete removal of ethanol is necessary for the RNA to perform well in downstream applications.**
- Air dry the RNA pellet for 3–5 min.
  - Do not completely dry the RNA pellet as this will greatly decrease its solubility. 
- Dissolve RNA in the buffer of your choice.
	- Dissolve RNA by passing the solution a few times through a pipette tip or by vigorous vortexing
- The resuspension volume is determined by the size of the RNA pellet. 3–5 mm pellets typically require 300–500 μL. If necessary, increase the resuspension volume or incubate at 55–60°C to completely dissolve the pellet.
- Store at 4°C for immediate analysis. For long-term storage, store at –70°C or colder.
