---
title: "Synthesis EecSeq"
output: html_notebook
params:
  num_samples: 4
---
```{r setup}
library(tidyverse)
```

```{r}
synthesis <- tribble(~item, ~quantity,
                     "tips_10", 2 * params$num_samples,
                     "tips_100", 4 * params$num_samples,
                     "tubes_200ul", 1 * params$num_samples,
                     "kapa_stranded_mrna_kit_script", 2 * params$num_samples,
                     "kapa_first_synthesis_buffer_ul", 22 * params$num_samples,
                     "kapa_second_strand_marking_ul", 62 * params$num_samples, 
                     "kapa_second_enzyme_ul", 4 * params$num_samples)
```


---

- Place the plate/tubes in a thermal cycler and carry out the fragmentation and priming program as follows:

|Desired Fragment Size| Temp.| Duration|
|---------------------|------|---------|
|100 – 200 bp|94 °C|8 min|
|200 – 300 bp|94 °C|6 min|
|300 – 400 bp|85 °C|6 min|

##### For Puritz and Lotterhos (2017), we chose 94 °C for 7 mins to have fragments between 150-250 bp, approximately the same size distribution as planned for our DNA libraries.

- Immediately place the plate/tube on a magnet to capture the beads, and incubate until the liquid is clear. **Caution: To prevent hybridization of poly(A)- rich RNA to the capture beads, do not allow the sample to cool before placing on the magnet**-
- Carefully remove 10 μl of the supernatant containing the eluted, fragmented, and primed RNA into a separate plate or tube.
- Proceed immediately to **1st Strand Synthesis**.

### 1st Strand Synthesis
- On ice, assemble the 1st Strand Synthesis reaction as follows:

|Component|Volume|
|---------|------|
|Fragmented, primed RNA eluted from beads| 10 μl|
|1st Strand Synthesis Master Mix| 5 μl|
|--Total Volume--| --15 μl--|

- Keeping the plate/tube on ice, mix thoroughly by gently pipetting the reaction up and down several times.
- Incubate the plate/tube using the following protocol:

|Step|Temp.|Duration|
|----|-----|--------|
|Primer extension|25 °C|10 min|
|1st Strand synthesis|42 °C|15 min|
|Enzyme inactivation|70 °C|15 min|
|HOLD|4 °C|∞|

- Place the plate/tube on ice and proceed immediately to --2nd Strand Synthesis and Marking--.

### 2nd Strand Synthesis and Marking
- Assemble the 2nd Strand Synthesis and Marking reaction as follows:

|Component|Volume|
|---------|------|
|1st Strand cDNA| 15 μl|
|2nd Strand Synthesis and Marking Master Mix| 15 μl|
|--Total Volume--| --30 μl--|

- Mix thoroughly by gently pipetting the reaction up and down several times.
- Incubate the plate/tube using the following protocol:

|Step|Temp.|Duration|
|----|-----|--------|
|2nd Strand synthesis and marking|16 °C|60 min|
|HOLD|4 °C|∞|

- Place the plate/tube on ice and proceed immediately to --2nd Strand Synthesis and Marking Cleanup--.
