---
title: "Working with adapters"
output: 
  html_notebook: 
    code_folding: hide
---
## To order oligos:
- Go to [IDT website](www.idtdna.com), log in with the usualy pinsky credentials (MRS has in contacts) and view past orders. Click on the order and then hit the plus sign to see which oligos were purchased. Hit edit to update or change the oligo (like dropping the biotinylation from P2.2 or adding a phosphotioated bonds).  
- To order new oligos that have never been used before, sign in and click on custom DNA or RNA oligos, click on "Order Now" which should take you to an "Oligo Entry" page, and copy and paste your sequence, unless you have many, in which case you can click on "Bulk Input" to upload a file.  Then click "add to order".  

*Double check the guaranteed nmole that you will be getting back.*  

## To rehydrate oligos:
*Resuspend Oligos*  
IDT says to multiply # of nmoles on vial by 10 and add the answer uL of buffer to bring concentration to 100uM.  We want 200uM so we are multiplying by 5.

9 x 10^-9mol  = 0.045 x 10^-3 L = 45uL of buffer to resuspend
200 x 10^-6mol

9 x 5 = 45

- Go to the google doc ["Oligo”](https://docs.google.com/spreadsheets/d/1K2FYSjhMS82SeHkQGW1C9GFlu9ZZ_iZJVDfYxMP3SIE/edit?usp=sharing) in the Pinsky Lab folder.  
- Click on the “Resuspension” Tab.  
- Scroll to the bottom, enter the date of resuspension, name of adapter on tube, nmol of adapter on tube, and drag down the “amount of TE…” and “Final mol” cells

Spun dried tubes 30 sec @ 20,000g
Add TE buffer
Vortex 15 sec
Incubate at 4˚C overnight (on 11/24 we incubated for 1 hour)

## TE buffer recipe from IDT
lower pH with HCl or increase with NaOH.
```{r}
readr::read_csv("item, intial_mM, initial_vol_uL, final_vol_ul, final_mM
         Tris-HCl pH8, 1000, 500, 50, 10
         EDTA ph8, 500, 10, 50, 100")
```
fill to 50mL w/ distilled water (dH2O)

## 10x annealing buffer recipe - 50mL
```{r}
readr::read_csv("item, intial_mM, initial_vol_mL, final_vol_ml, final_mM
         Tris-HCl pH8, 1000, 5, 50, 100
         EDTA ph8, 500, 1, 50, 10
         NaCl, 5000, 5, 50, 500 ")
```
fill to 50mL w/ dH2O

## To anneal adapters  
To make 200uL of 40uM adapters  

  - 40uL 1.1 adapter 200uM  
  - 40uL 1.2 adapter 200uM  
  - 20uL of 10x annealing buffer  
  - 100uL pH2O 
  
incubate @ 97.5˚C 2.5 min; cool @ 3˚ per minute to 21˚C, hold @ 4˚C   


