---
title: "RNA extraction for EecSeq"
date: "`r Sys.Date()`"
output: html_notebook
params:
  num_samples: 1
  mass_sample_mg: 15
  mulitplier_tri: 15
---

Explanation of params:  
- num_samples: the number of samples included in this extraction.  We combined 2 fish into one vial.  
- mass_sample_mg: the total mass of all fish included in each vial.  Total mass was 15mg.  
- multiplier_tri: number of volumes to multiple mass by to determine volume of tri-reagent to use.  The protocol said to use 10-20 volumes, so we used 15.

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
```

Manufacturers instructions can be found [here](https://github.com/pinskylab/laboratory/blob/master/documentation/tri-reagent-protocol.pdf).

## RNA Prep
```{r echo=FALSE}
rna_extr <- tribble(~item, ~quantity,
                    "tri_reagent_ul", 1000,
                    "tubes_1.7ml", 3,
                    "bcp_ul", 100,
                    "tips_1000", 7,
                    "tips_100", 2,
                    "isopropanol_ul", 500, 
                    "ethanol_ul", 1000,
                    "1x_te_buffer_ul", 300
                    ) %>% 
  mutate(quantity = quantity* params$num_samples)
```

```{r}
vol_tri_reagent_ul = params$mulitplier_tri * params$mass_sample_mg
vol_bcp_ul = vol_tri_reagent_ul * .1
vol_iso_ul = vol_tri_reagent_ul * .5
```


### Extract RNA from individuals to be used for probes

#### Reagents and supplies
- Nuclease-free Water
- 1-bromo-3-chloropropane (BCP; recommended, e.g., MRC, Cat #BP 151), or chloroform without added isoamyl alcohol 100% 
- isopropanol, ACS grade or better
- 100% ethanol, ACS grade or better

#### Equipment
- Pestles for 1.5mL tubes
- Appropriately sized RNase-free centrifuge tubes with secure closures, compatible with phenol/chloroform (polypropyl- ene, or polyallomer), and capable of withstanding centrifugal forces of 12,000G
- Centrifuge capable of 12,000 x g and 4˚C - we don't have that.  Because we tried the large centrifuge at cold temps and slow speeds (3220 x g) and it didn't work, we are trying placing the small centrifuge in a plastic tub the snow with another plastic tub as the lid, high today outside is 3˚C.  

**Notes before starting**
- Place all pipettes, tips, and supplies inside the hood and expose to UV for 30 minutes. (Do not put your samples in the UV hood!)  
- Prepare 75% ethanol by mixing 250 μL nuclease-free water with 750 μL 100% ethanol per mL of TRI Reagent solution to be used. Include 10% overage to ensure a sufficient volume.

#### Procedure
1. Homogenize tissue samples in `r vol_tri_reagent_ul` ul TRI Reagent solution.  
2. Incubate the homogenate for 5 min at room temp.  
3. Centrifuge at 12,000 x g for 10 min at 4°C and transfer the supernatant to a fresh tube.  
4.    a. Add `r vol_bcp_ul` μL BCP.  
      b. Cap the tubes tightly and shake vigorously for 15 sec.  
      c. Incubate the mixture at room temperature for 5–15 min.  
      
5.    a. Centrifuge at 12,000 x g for 10–15 min at 4°C, then transfer the aqueous phase to a fresh tube. *Centrifugation at temperatures > 8°C may cause some DNA to partition in the aqueous phase, which is undesireable.*  
      b. Transfer the aqueous phase (colorless top layer) to a fresh tube.
6. Add `r vol_iso_ul` μL of isopropanol, vortex for 5–10 sec, and incubate at room temp for 5–10 min.  
      
7.    a. Centrifuge at 12,000 x g for 8 min at 4–25°C, and discard the supernatant.  
      b. Carefully remove the supernatant without disturbing the pellet.  *Precipitated RNA forms a gel-like or white pellet on the side and bottom of the tube.*
      
8. Add 1 mL of 75% ethanol per 1 mL of TRI Reagent solution. 

9.    a. Centrifuge at 7,500 x g for 5 min at 4–25°C. *If the precipitated RNA floats or does not form a compact pellet, repeat the centrifugation at 12,000 x g for 5 min to consolidate the pellet at the bottom of the tube.*  
      b. Remove the ethanol wash without disturbing the pellet.  
      c. Remove all residual ethanol by centrifuging again briefly and removing the ethanol that collects with a fine tip pipette. **Complete removal of ethanol is necessary for the RNA to perform well in downstream applications.**  
      d. Air dry the RNA pellet for 3–5 min.  Do not completely dry the RNA pellet as this will greatly decrease its solubility.  
10. a. Dissolve RNA in the buffer of your choice.
	- Dissolve RNA by passing the solution a few times through a pipette tip or by vigorous vortexing
- The resuspension volume is determined by the size of the RNA pellet. 3–5 mm pellets typically require 300–500 μL. If necessary, increase the resuspension volume or incubate at 55–60°C to completely dissolve the pellet.
- Store at 4°C for immediate analysis. For long-term storage, store at –70°C or colder.
